National demographic-disturbance relationships and local monitoring both reduce uncertainty in projections of boreal caribou population viability: tools to support monitoring decisions

### 2023-08-21 analysis/paper/paper.Rmd
prior calibration

runScenarios = F 
doPriorCalibration=T #Note the prior calibration is time consuming. 
```

@@

  #TO DO: move CaribouDemo_v1.15Reduce.R to paper directory. And produce pngs, rather than pdfs.
  source(here::here("analysis/scripts/calibratePriors.R"))
}else{
  KS<-read.csv(here::here("tabs/SurvKSAll.csv"))
  KSRec<-read.csv(here::here("tabs/RecKSAll.csv"))

}
  png(here::here("figs/PriorKSDists.png"),
    height = 4, width = fig_widths["two"], units = "in",res=600)
    base = ggplot(KS,aes(x=lse,y=KSDistance,color=sse))+
    geom_point()+facet_grid(scn~ssv,labeller="label_both",scales="free_y")+theme_bw()+ylab("Kolmogorov-Smirnov Distance")+xlab("lse: multiplier on SD of survival intercept")
    print(base)
  dev.off()

  png(here::here("figs/PriorKSDistsRec.png"),
    height = 4, width = fig_widths["two"], units = "in",res=600)
    base = ggplot(KSRec,aes(x=lre,y=KSDistance,color=sre))+
    geom_point()+facet_grid(scn~srv,labeller="label_both",scales="free_y")+theme_bw()+ylab("Kolmogorov-Smirnov Distance")+xlab("lre: multiplier on SD of recruitment intercept")
    print(base)
  dev.off()
```

```{r PriorKSDists, fig.cap="Prior parameter values should be set so that, given minimal monitoring data ('All' scenario: 1 year of 1 collar per year), the posterior survival distribution is similar to the distribution of simulated outcomes from the national model, measured here by Komogorov-Smirnov distance [ks.test function in @R_R_2022]. Priors should also be vague enough so that, given intensive monitoring of populations where survival is higher or lower than the national average ('High' and 'Low' scenarios: 15 years of 45 collars per year from 97.5% and 2.5% quantiles of the national model), posteriors reflect local survival, rather than priors from the national model. When the multiplier on the prior standard deviation of the survival intercept (lse: $\\beta^S_0$ in Table @ref(tab:priors)) is 5, the prior is informative enough to meet the former condition, and vague enough to meet the latter. Given lse=5, low values of the mean standard deviations of the prior distribution of the random effect of year on survival (sse=0.034784, ssv=0.01, $\\sigma^2_S$ in Table @ref(tab:priors)) yield priors that meet both conditions (Figs @ref(fig:priorAllFits) and @ref(fig:priorLowFits)). Anthropogenic disturbance is set to 0 in these scenarios." , cache=TRUE}
knitr::include_graphics(here::here("figs/PriorKSDists.png"))

@@

#Note: file wouldn't knit for Anna without access to all the newest png files below, so just commenting out the problem ones for now
```

```{r PriorAllFits, fig.cap="Differences between posterior predictive survival distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming minimal monitoring (1 years of 1 collar per year) of populations from the 50% quantile of the national model. In these examples, the standard deviation the random effect prior (ssv: $\\sigma^2_S$ in Table @ref(tab:priors)) is 0.03, and  anthropogenic disturbance is 0. Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.",  cache=TRUE}
# COMMENTED OUT FOR NOW CAUSE I DIDN"T HAVE ACCESS
knitr::include_graphics(here::here("figs/SurvsQStarts1ssv0.03.png"))
```

```{r PriorLowFits, fig.cap="Posterior predictive survival distributions from the Bayesian model ('local') assuming intensive monitoring (15 years of 45 collars per year) of populations from the 2.5% quantile of the national model. In these examples, the standard deviation the random effect prior (ssv: $\\sigma^2_S$ in Table @ref(tab:priors)) is 0.03, and  anthropogenic disturbance is 0. Bands show 95% posterior predictive intervals, black dots show true survival rates from the simulated population, and open triangles show observed survival rates.", cache=TRUE}
# COMMENTED OUT FOR NOW CAUSE I DIDN"T HAVE ACCESS
knitr::include_graphics(here::here("figs/SurvsQStarts45lowssv0.03.png"))
```

```{r PriorAllFitsAnthro90, fig.cap="Differences between posterior predictive survival distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming minimal monitoring (1 years of 1 collar per year) of populations from the 50% quantile of the national model. In these examples, anthropogenic disturbance is 90%. When the multiplier on the prior standard deviation of the survival slope (bse: $\\beta^S_a$ in Table @ref(tab:priors)) is 3, the posterior survival distribution is similar to the distribution of simulated outcomes from the national model. Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.", cache=TRUE}
knitr::include_graphics(here::here("figs/SurvsQStarts1Anthro90.png"))
```

```{r PriorHighFitsAnthro90, fig.cap="Posterior predictive survival distributions from the Bayesian model, assuming intensive monitoring (15 years of 45 collars per year) of populations from the 97.5% quantile of the national model. In these examples, anthropogenic disturbance is 90%, and the effect of anthropogenic disturbance varies from none (sS=0) through expected (sS=1) to double the expected slope (sS=2). All considered multipliers on the prior standard deviation of the survival slope (bse: $\\beta^S_a$ in Table @ref(tab:priors)) are sufficiently vague to reflect local variation among populations. Bands show 95% posterior predictive intervals, black dots show true survival rates from the simulated population, and open triangles show observed survival rates.", cache=TRUE}
knitr::include_graphics(here::here("figs/SurvsQStarts45Anthro90high.png"))
```

```{r PriorKSDistsRec, fig.cap="Prior parameter values should be set so that, given no monitoring data ('All' scenario), the posterior recruitment distribution is similar to the distribution of simulated outcomes from the national model, measured here by Komogorov-Smirnov distance [ks.test function in @R_R_2022]. Setting the multiplier on the prior standard deviation of the recruitment intercept (lre: $\\beta^R_0$ in Table @ref(tab:priors)) to 3, and values of the mean and standard deviations of the prior distribution of the random effect of year on recruitment to sre=0.23 and srv=0.22, respectively ($\\sigma^2_R$ in Table @ref(tab:priors)) allows us to meet this condition (Fig @ref(fig:priorAllFitsRec)). Priors should also be vague enough so that, given intensive monitoring of populations where recruitment is higher or lower than the national average ('High' and 'Low' scenarios: 15 years of 135 cows per year from 97.5% and 2.5% quantiles of the national model), posteriors reflect local recruitment, rather than priors from the national model. When the multiplier on the prior standard deviation of the recruitment intercept (lre: $\\beta^R_0$ in Table @ref(tab:priors)) is 1, the prior is not sufficiently vague to meet the latter condition (Fig @ref(fig:PriorLowFitsRec)). Anthropogenic disturbance is set to 0 in these scenarios." , cache=TRUE}
knitr::include_graphics(here::here("figs/PriorKSDistsRec.png"))
```

```{r PriorAllFitsRec, fig.cap="Differences between posterior predictive recruitment distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming no monitoring. In these examples, the standard deviation the random effect prior (srv: $\\sigma^2_R$ in Table @ref(tab:priors)) is 0.22, and  anthropogenic disturbance is 0. Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.",  cache=TRUE}
knitr::include_graphics(here::here("figs/RecrQStarts1srv0.22.png"))
```

```{r PriorLowFitsRec, fig.cap="Posterior predictive recruitment distributions from the Bayesian model ('local') assuming intensive monitoring (135 cows per year for 15 years) of populations from the 2.5% quantile of the national model.  In these examples, the standard deviation the random effect prior (srv: $\\sigma^2_R$ in Table @ref(tab:priors)) is 0.22, and  anthropogenic disturbance is 0. Bands show 95% posterior predictive intervals, black dots show true recruitment rates from the simulated population, and open triangles show observed recruitment rates.", cache=TRUE}
knitr::include_graphics(here::here("figs/RecrQStarts45lowsrv0.22.png"))
```

```{r PriorAllFitsAnthro90Rec, fig.cap="Differences between posterior recruitment predictive distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming no monitoring. In these examples, anthropogenic disturbance is 90%. When the multiplier on the prior standard deviation of the slope of the relationship between recruitment and anthropogenic disturbance (bre: $\\beta^R_a$ in Table @ref(tab:priors)) is 4, the posterior recruitment distribution is similar to the distribution of simulated outcomes from the national model. Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.", cache=TRUE}
knitr::include_graphics(here::here("figs/RecrQStarts1Anthro90.png"))
```

```{r PriorHighFitsAnthro90Rec, fig.cap="Posterior predictive recruitment distributions from the Bayesian model, assuming intensive monitoring (15 years of 135 collars per year) of populations from the 97.5% quantile of the national model. In these examples, anthropogenic disturbance is 90%, and the effect of anthropogenic disturbance varies from none (rS=0) through expected (rS=1) to double the expected slope (rS=2). All multipliers on the prior standard deviation of the survival slope (bre: $\\beta^R_a$ in Table @ref(tab:priors)) are sufficiently vague to reflect local variation among populations. Bands show 95% posterior predictive intervals, black dots show true recruitment rates from the simulated population, and open triangles show observed recruitment rates.", cache=TRUE}
knitr::include_graphics(here::here("figs/RecrQStarts45Anthro90high.png"))
```

>>>>

runScenarios = F 
doPriorCalibration=F #Note the prior calibration is time consuming. 
```

@@

  source(here::here("analysis/scripts/calibratePriors.R"))
}else{
  KS<-read.csv(here::here("tabs/SurvKSAll.csv"))
  KSRec<-read.csv(here::here("tabs/RecKSAll.csv"))

}
  #library(ggplot2);fig_widths=data.frame(two=7)
  KS$SigmaMean = KS$sInterannualVar
  KS$SigmaSD = KS$sInterannualVarSE
  KS=subset(KS,scn=="All")
  png(here::here("figs/PriorKSDists.png"),
    height = 3, width = fig_widths["two"], units = "in",res=100)
    base = ggplot(KS,aes(x=sIntSEMod,y=KSDistance,color=SigmaMean))+
    geom_point()+facet_wrap(~SigmaSD,labeller="label_both")+theme_bw()+ylab("Kolmogorov-Smirnov Distance")+xlab("sIntSEMod: multiplier on SD of survival intercept")
    print(base)
  dev.off()

  KSRec$SigmaMean = KSRec$rInterannualVar
  KSRec$SigmaSD = KSRec$rInterannualVarSE
  KSRec=subset(KSRec,scn=="All")
  png(here::here("figs/PriorKSDistsRec.png"),
    height = 3, width = fig_widths["two"], units = "in",res=600)
    base = ggplot(KSRec,aes(x=rIntSEMod,y=KSDistance,color=SigmaMean))+
    geom_point()+facet_wrap(~SigmaSD,labeller="label_both")+theme_bw()+ylab("Kolmogorov-Smirnov Distance")+xlab("rIntSEMod: multiplier on SD of recruitment intercept")
    print(base)
  dev.off()
```

Prior parameter values should be set so that, given no monitoring data, the prior survival distribution is similar to the distribution of simulated outcomes from the national model in the case of no anthropogenic disturbance. This condition is met when the multiplier on the prior standard deviation of the survival intercept (sIntSEMod: $\\beta^S_0$ in Table @ref(tab:priors)) is 5 and the mean and standard deviations of the prior distribution of the random effect of year on survival are low  (SigmaMean & SigmaSD, $\\sigma^2_S$ in Table @ref(tab:priors)) (Figs @ref(fig:priorKSDists) and @ref(fig:priorAllFits)). Priors should also be vague enough so that, given intensive monitoring of populations where survival is higher or lower than the national average ('High' and 'Low' scenarios: 15 years of 60 collars per year from 97.5% and 2.5% quantiles of the national model), posteriors reflect local survival, rather than priors from the national model. Setting the mean and standard deviations of the prior distribution of the random effect of year on survival to moderately low values (SigmaMean=0.034784, SigmaSD=0.03, $\\sigma^2_S$ in Table @ref(tab:priors)) allows both conditions to be met (Figs @ref(fig:priorKSDists), @ref(fig:priorAllFits) and @ref(fig:priorLowFits)).

```{r PriorKSDists, fig.cap="Similarity of the prior survival distribution to the distribution of simulated outcomes from the national model measured by Kolmogorov-Smirnov distance [ks.test function in @R_R_2022]. The distributions are similar when the multiplier on the prior standard deviation of the survival intercept (sIntSEMod: $\\beta^S_0$ in Table @ref(tab:priors)) is 5 and the mean and standard deviations of the prior distribution of the random effect of year on survival are low  (SigmaMean & SigmaSD, $\\sigma^2_S$ in Table @ref(tab:priors)). Anthropogenic disturbance is set to 0 in these scenarios." , cache=TRUE}
knitr::include_graphics(here::here("figs/PriorKSDists.png"))

@@

```{r PriorAllFits, fig.cap="Differences between prior predictive survival distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming no local monitoring. In these examples, the standard deviation of the random effect prior (SigmaSD: $\\sigma^2_S$ in Table @ref(tab:priors)) is 0.03, and anthropogenic disturbance is 0. Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.",  cache=TRUE}

knitr::include_graphics(here::here("figs/SurvInterceptAllsInterannualVarSE0.03.png"))
```

```{r PriorLowFits, fig.cap="Posterior predictive survival distributions from the Bayesian model ('local') assuming intensive monitoring (15 years of 60 collars per year) of populations from the 2.5 \\% quantile of the national model. In these examples, the standard deviation of the random effect prior (SigmaSD: $\\sigma^2_S$ in Table @ref(tab:priors)) is 0.03, and  anthropogenic disturbance is 0. Bands show 95 \\% posterior predictive intervals, black dots show true survival rates from the simulated population, and open triangles show observed survival rates. When the multiplier on the prior standard deviation of the survival intercept (sIntSEMod: $\\beta^S_0$ in Table @ref(tab:priors)) is too low (<5) the insufficiently vague priors substantially influence the posterior distributions, yielding poor local parameter estimates.", cache=TRUE}

knitr::include_graphics(here::here("figs/SurvsurvInterceptLowsInterannualVarSE0.03.png"))
```

```{r PriorAllFitsAnthro90, fig.cap="Differences between prior predictive survival distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming no local monitoring. In these examples, anthropogenic disturbance is 90\\%. When the multiplier on the prior standard deviation of the survival slope (sSlopeSEMod: $\\beta^S_a$ in Table @ref(tab:priors)) is 3, the posterior survival distribution is similar to the distribution of simulated outcomes from the national model. Bands show 95\\% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.", cache=TRUE}
knitr::include_graphics(here::here("figs/SurvSlopeAll.png"))
```

```{r PriorHighFitsAnthro90, fig.cap="Posterior predictive survival distributions from the Bayesian model, assuming intensive monitoring (15 years of 60 collars per year) of populations from the 97.5 \\% quantile of the national model. In these examples, anthropogenic disturbance is 90\\%, and the effect of anthropogenic disturbance varies from none (sSlopeMod=0) through expected (sSlopeMod=1) to double the expected slope (sSlopeMod=2). All considered multipliers on the prior standard deviation of the survival slope (sSlopeSEMod: $\\beta^S_a$ in Table @ref(tab:priors)) are sufficiently vague to reflect local variation among populations. Bands show 95 \\% posterior predictive intervals, black dots show true survival rates from the simulated population, and open triangles show observed survival rates.", cache=TRUE}
knitr::include_graphics(here::here("figs/SurvSlopeHigh.png"))
```

Similarly, prior parameter values should be set so that, given no monitoring data, the prior recruitment distribution is similar to the distribution of simulated outcomes from the national model in the case of no anthropogenic disturbance. This condition is met when the multiplier on the prior standard deviation of the recruitment intercept (rIntSEMod: $\\beta^R_0$ in Table @ref(tab:priors)) is 4 and the mean and standard deviations of the prior distribution of the random effect of year on survival are low  (SigmaMean & SigmaSD, $\\sigma^2_R$ in Table @ref(tab:priors)) (Figs @ref(fig:priorKSDistsRec) and @ref(fig:priorAllFitsRec)). Priors should also be vague enough so that, given intensive monitoring of populations where recruitment is higher or lower than the national average ('High' and 'Low' scenarios: 15 years of 9*60 cows per year from 97.5% and 2.5% quantiles of the national model), posteriors reflect local recruitment, rather than priors from the national model. Setting the mean and standard deviations of the prior distribution of the random effect of year on recruitment to moderately low values (SigmaMean=0.23, SigmaSD=0.22, $\\sigma^2_R$ in Table @ref(tab:priors)) allows both conditions to be met (Figs @ref(fig:priorKSDistsRec), @ref(fig:priorAllFitsRec) and @ref(fig:priorLowFitsRec)).

```{r PriorKSDistsRec, fig.cap="Similarity of the prior recruitment distribution to the distribution of simulated outcomes from the national model measured by Kolmogorov-Smirnov distance [ks.test function in @R_R_2022]. The distributions are similar when the multiplier on the prior standard deviation of the survival intercept (rIntSEMod: $\\beta^R_0$ in Table @ref(tab:priors)) is 4 and the mean and standard deviations of the prior distribution of the random effect of year on survival are low  (SigmaMean & SigmaSD, $\\sigma^2_R$ in Table @ref(tab:priors)). Anthropogenic disturbance is set to 0 in these scenarios." , cache=TRUE}
knitr::include_graphics(here::here("figs/PriorKSDistsRec.png"))
```

```{r PriorAllFitsRec, fig.cap="Differences between prior predictive recruitment distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming no local monitoring. In these examples, the standard deviation of the random effect prior (SigmaSD: $\\sigma^2_R$ in Table @ref(tab:priors)) is 0.22, and anthropogenic disturbance is 0. Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.",  cache=TRUE}
knitr::include_graphics(here::here("figs/RecInterceptAllrInterannualVarSE0.22.png"))
```

```{r PriorHighFitsRec, fig.cap="Posterior predictive recruitment distributions from the Bayesian model ('local') assuming intensive monitoring (9*60 cows per year for 15 years) of populations from the 97.5% quantile of the national model.  In these examples, the standard deviation the random effect prior (SigmaSD: $\\sigma^2_R$ in Table @ref(tab:priors)) is 0.22, and  anthropogenic disturbance is 0. Bands show 95% posterior predictive intervals, black dots show true recruitment rates from the simulated population, and open triangles show observed recruitment rates.", cache=TRUE}
knitr::include_graphics(here::here("figs/RecInterceptHighrInterannualVarSE0.22.png"))
```

```{r PriorAllFitsAnthro90Rec, fig.cap="Differences between prior predictive survival distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming no local monitoring. In these examples, anthropogenic disturbance is 90\\%. When the multiplier on the prior standard deviation of the recruitment slope (rSlopeSEMod: $\\beta^R_a$ in Table @ref(tab:priors)) is 4, the posterior survival distribution is similar to the distribution of simulated outcomes from the national model. Bands show 95\\% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.", cache=TRUE}
knitr::include_graphics(here::here("figs/RecSlopeAll.png"))
```

```{r PriorHighFitsAnthro90Rec, fig.cap="Posterior predictive recruitment distributions from the Bayesian model, assuming intensive monitoring (15 years of 60 collars per year) of populations from the 97.5 \\% quantile of the national model. In these examples, anthropogenic disturbance is 90\\%, and the effect of anthropogenic disturbance varies from none (rSlopeMod=0) through expected (rSlopeMod=1) to double the expected slope (rSlopeMod=2). All considered multipliers on the prior standard deviation of the recruitment slope >1 (rSlopeSEMod: $\\beta^R_a$ in Table @ref(tab:priors)) are sufficiently vague to reflect local variation among populations. Bands show 95 \\% posterior predictive intervals, black dots show true survival rates from the simulated population, and open triangles show observed survival rates.", cache=TRUE}
knitr::include_graphics(here::here("figs/RecSlopeHigh.png"))
```
